---
title: "Zillow Project Report"
author: "Liz Vincent"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message = F, warning = F,fig.align = "center")

library(data.table)
library(corrplot)
library(dplyr)
library(ggplot2)
library(magrittr)
library(stringi)
library(stringr)
library(rebus)
library(gridExtra)
library(reshape2)
library(ggmap)

properties<-read.csv("properties_2016.csv",stringsAsFactors = F)  # This takes a while...~3 million rows, 58 columns
train<-read.csv("train_2016_v2.csv")
```

```{r train_variable_class}
train<-merge(train,properties,by="parcelid")
factor_vars <- str_subset(names(train),pattern="id")
factor_vars <- c(factor_vars,"fips","propertycountylandusecode","propertyzoningdesc","rawcensustractandblock","censustractandblock")
train[,factor_vars]<-lapply(train[,factor_vars],factor)
```

```{r train_missingness, include=FALSE}
# Count the number of NAs or empty strings in each column of properties_2016
missingness<-lapply(train,function(x){
    ifelse(is.character(x),
          sum(str_detect(x,pattern=START %R% END)),
          sum(is.na(x))
  )
})

# Format missingness
missingness<-round(unlist(missingness)/nrow(train)*100,2)
missingness<-sort(missingness,decreasing=T)
missingness<-as.data.frame(missingness)
missingness$var<-stri_trans_totitle(row.names(missingness))
row.names(missingness)<-seq(1:nrow(missingness))
```

```{r, fig.height=6, fig.width=8}
coordinates<-train[,c("longitude","latitude")]

coordinates$latitude<-as.numeric(str_replace(coordinates$latitude, pattern=capture(dgt(2)) %R% capture(one_or_more(DGT)),replacement=REF1 %R% DOT %R% REF2))
coordinates$longitude<-as.numeric(str_replace(coordinates$longitude, pattern=capture("-" %R% dgt(3)) %R% capture(one_or_more(DGT)),replacement=REF1 %R% DOT %R% REF2))


get_density <- function(x, y, n = 100) {
  dens <- MASS::kde2d(x = x, y = y, n = n)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
#cali <- c(left = -125, bottom = 30, right = -115, top = 40)
map <- get_googlemap(center=c(lon = -118.4, lat=34), zoom = 8, maptype = "roadmap")
ggmap(map) +
  geom_point(aes(x = longitude, y = latitude,color=get_density(x=longitude,y=latitude)), data = coordinates) +
  scale_color_gradientn("Density",colors = c("blue4","blue","springgreen2","yellow","orange","red","red3"), breaks=c(0.1,3),labels=c("low","high")) +
  labs(x="Longitude",y="Latitude")
```

# 1. Tidy the Data

# 2. Address Missingness

# 3. Impute Missing Values
Although at first glance it appears that many values have very high missingness, there are several instances in which "missing" data may actually be a negative response. For example, `taxdelinquencyflag` is missing for `r missingness[missingness$var == 'Taxdelinquencyflag','missingness']`% of observations, when in reality these are probably observations in which there has not been a tax delinquency flag. Therefore, I will not exclude tax delinquency tag or year based on missingness, but will impute the missing values for tax delinquency tag as negative responses.