---
title: "Zillow Project Report"
author: "Liz Vincent"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message = F, warning = F,fig.align = "center")

packages<-c("data.table","corrplot","dplyr","ggplot2","magrittr","stringi","stringr","rebus","gridExtra","reshape2","ggmap","tidyr","caret","xgboost","Matrix","ranger")

for (i in packages){
  if(!require(i,character.only = T,quietly=T,warn.conflicts = F)){
    install.packages(i)
  }
  require(i,character.only = T,quietly=T,warn.conflicts = F)
}

properties_2016<-read.csv("properties_2016.csv",stringsAsFactors = F)  # This takes a while...~3 million rows, 58 columns
train_2016<-read.csv("train_2016_v2.csv")
```

```{r train_variable_class}
train_2016<-merge(train_2016,properties,by="parcelid")
factor_vars <- str_subset(names(train_2016),pattern="id")
factor_vars <- c(factor_vars,"fips","propertycountylandusecode","propertyzoningdesc","rawcensustractandblock","censustractandblock")
train_2016[,factor_vars]<-lapply(train_2016[,factor_vars],factor)
```

```{r train_missingness, echo=F}
# Count the number of NAs or empty strings in each column of properties_2016
missingness_2016_preimputation<-lapply(train_2016,function(x){
    ifelse(is.character(x),
          sum(str_detect(x,pattern=START %R% END)),
          sum(is.na(x))
  )
})

# Format missingness
missingness_2016_preimputation<-round(unlist(missingness_2016_preimputation)/nrow(train_2016)*100,3)
missingness_2016_preimputation<-sort(missingness_2016_preimputation,decreasing=T)
missingness_2016_preimputation<-as.data.frame(missingness_2016_preimputation)
missingness_2016_preimputation$var<-row.names(missingness_2016_preimputation)
row.names(missingness_2016_preimputation)<-seq(1:nrow(missingness_2016_preimputation))
missingness_2016_preimputation<-select(missingness_2016_preimputation,c("var","missingness_2016_preimputation"))
missingness_2016_preimputation<-missingness_2016_preimputation %>% mutate(category = lapply(missingness_2016_preimputation,function(x){if(x < 5){"<5%"} else if(x>=5 & x <= 95){">=5%, <=95%"} else{">95%"}}))
missingness_2016_preimputation$category<-as.character(missingness_2016_preimputation$category)

labels<-c("Framing Type","Perimeter Living Area Sqft.","Type of Floors in Multi-Story House","Basement Sqft.","Yard Building Sqft.","Fireplace, Y/N","Architecture Style","Construction Material","Unfinished and Finished Sqft.","Type of Deck","Pool Sqft.","Spa or Hot Tub, Y/N","Pool with Spa or Hot Tub, Y/N", "Year of Tax Delinquency", "Tax Delinquency, Y/N","Spa or Hot Tub, Y/N","Patio Sqft.","Total Sqft.","First Floor Living Area Sqft.","First Floor Living Area Sqft.","No. Fireplaces","No. 3/4 Baths","Pool without Spa or Hot Tub, Y/N","No. Pools","No. Stories","Type of Air Conditioning","Garage Sqft.","No. Garages","Neighborhood ID","Type of Heating System","Building Condition","No. Units","Lot Sqft.","Living Area Sqft.","City ID","No. Full Baths","Total No. Baths","Year Built","Total Finished Living Area Sqft.","Census Tract and Block ID","Assessed Value of Structure","Zip Code","Total Property Tax","Total Tax Assessed Value of Parcel","Assessed Value of Land","Transaction Date","Total No. Rooms","County ID","Census Tract and Block ID","Description of Zoning","Zoning","Zoning at County Level","Parcel ID","Longitude","Log Error","Latitude", "Federal Information Processing Standard Code","No. Bedrooms","Total No. Baths","Asessment Year")

p1 <- ggplot(missingness_2016_preimputation,aes(x=reorder(var, missingness_2016_preimputation),y=missingness_2016_preimputation,fill=category)) +
  geom_col() +
  scale_fill_manual(breaks=c(">95%",">=5%, <=95%","<5%"),values=c("blue","black","red")) +
  scale_y_continuous(limits = c(NA,100),expand = c(0,0),breaks=c(0,25,50,75,99), labels=c("0","25","50","75","100")) +
  theme_bw() +
  theme(panel.grid = element_blank(),legend.position = c(0.9,0.1), axis.ticks=element_blank()) +
  guides(fill=guide_legend(element_blank())) +
  labs(y="Missingness (%)",x=element_blank()) +
  scale_x_discrete(breaks=missingness_2016_preimputation$var, labels=labels) +
  coord_flip()

p1
```

```{r, fig.height=6, fig.width=8}
coordinates<-train[,c("longitude","latitude")]

coordinates$latitude<-as.numeric(str_replace(coordinates$latitude, pattern=capture(dgt(2)) %R% capture(one_or_more(DGT)),replacement=REF1 %R% DOT %R% REF2))
coordinates$longitude<-as.numeric(str_replace(coordinates$longitude, pattern=capture("-" %R% dgt(3)) %R% capture(one_or_more(DGT)),replacement=REF1 %R% DOT %R% REF2))


get_density <- function(x, y, n = 100) {
  dens <- MASS::kde2d(x = x, y = y, n = n)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
#cali <- c(left = -125, bottom = 30, right = -115, top = 40)
map <- get_googlemap(center=c(lon = -118.4, lat=34), zoom = 8, maptype = "roadmap")
ggmap(map) +
  geom_point(aes(x = longitude, y = latitude,color=get_density(x=longitude,y=latitude)), data = coordinates) +
  scale_color_gradientn("Density",colors = c("blue4","blue","springgreen2","yellow","orange","red","red3"), breaks=c(0.1,3),labels=c("low","high")) +
  theme(axis.text = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),legend.position = c(0.1,0.1),legend.background = element_blank())
```

# 1. Tidy the Data

# 2. Address Missingness

# 3. Impute Missing Values
Although at first glance it appears that many values have very high missingness, there are several instances in which "missing" data may actually be a negative response. For example, `taxdelinquencyflag` is missing for `r missingness[missingness$var == 'Taxdelinquencyflag','missingness']`% of observations, when in reality these are probably observations in which there has not been a tax delinquency flag. Therefore, I will not exclude tax delinquency tag or year based on missingness, but will impute the missing values for tax delinquency tag as negative responses.